---
title: "Delete files from list"
#output: html_notebook
---

# Build list containing files to be deleted
This uses the output of change_file_names.Rmd

Import the file produced by change_file_names.Rmd
```{r}
root = "/home/student12/project/"

# Load the file
sample_metadata = read_tsv(paste0(root,"data/old_new_filenames.txt"))

```
Filter out the files to be deleted using tidyverse and export to directory.

```{r}
library(tidyverse)

# Wish to remove the 3'_20 and 3'_30 mutants since not used in study
# use grepl to find the mutants
mutant_30 = grepl("3'_30", sample_metadata$new_file_names)
mutant_20 = grepl("3'_20", sample_metadata$new_file_names)

# Use dplyr's filter to produce a data.frame containing the files to be deleted
files_to_be_deleted = dplyr::filter(sample_metadata, mutant_30 | mutant_20)

# View files
head(files_to_be_deleted)
```

# Export file
Export the data.frame using write.table to the data directory. This is imortant since this file now keeps track of which files that have been deleted.
```{r}
write.table(files_to_be_deleted, paste0(root, "data/","files_to_be_deleted.txt"), sep="\t",row.names=FALSE)
```

# Delete files [Optional]
Delete the files contained in the exported data.frame.

```{r}
# Set working directory to directory containing files
path_to_files = setwd(paste0(root,"data/all_data/"))

# Uncomment if desired to delete, there is no way of recovering the deleted files
#file.remove(files_to_be_deleted[,1])

```

