---
title: "R Notebook"
#output: html_notebook
---

# Loading data
## Load data into R
```{r}
library(tidyverse)

```

```{r}
samples <- read_tsv("/home/student6/ProjectZIKA/team_project/data/Sample_table.txt", )
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# QC Preprocessing
Trim to minimal 35 bases and 35 quality score: **Trimmomatic v0.22**

# de novo assembly
**ABySS v1.3.7**


# Disassembly BAM files
**SamToFastq.jar** and **SAMtools**


# Alignment
(reference, de novo concensus) with very sensitive local alignment:
**Bowtie2 v2.2.4**



# Compress SAM file and sort by coordinate
**Samtool view** and **SortSam** using:
**SAMtools v1.1** and **Picard-tools v1.128**


# Remove PCR duplicates
**MarkDuplicates** and **Picard-tools v1.128**


# Variant detection
**V-phaser2 v2.0**


# Shannon entropy
**RF** and **bam2R** using:
**deepSNV v1.16.0**

```{r}
new_entropy = function(x){
  x[1:5]
  new_entropy = -sum(x*log(x), na.rm = TRUE)
  
  return(new_entropy)}
```


