---
title: "Simplifying Metadata"
#output: html_notebook
---

**This notebook simplifies the metadata file generated with the script data_acquisition.Rmd by removing the unwanted strains or passages for the study. This notebook can be used if you want to simplify the data set and focus on the study of one or several strains/passages. It generates a .txt file that will be used in the script delete_files_from_list.Rmd to delete the undesired files**

- **Input** : sample_metadata.txt 
- **Output** : simplified_metadata.txt 

```{r}
# Check workig directory
getwd()

# Define paths
root = "/home/student12/project/"
path_to_files = "data/all_data/"

# List the files in a directory
file_names = list.files(paste0(root,path_to_files), pattern = "*.fastq$")

# Look at list
head(file_names)
```


```{r}
library(tidyverse)

# Load metadata
sample_metadata <- read_tsv(paste0(root,"data/sample_metadata.txt"))

# Inspect metadata
head(sample_metadata)

```

```{r}
# Extract sample names from metadata
sample_metadata = cbind(sample_metadata, file_name = stringr::str_sub(sample_metadata$'Comment[FASTQ_URI]', -21, -4))

# Inspect changes
head(sample_metadata)

# Replace spaces with "_" and extract the necessary data
simplified_metadata = cbind(sample_metadata, new_file_names = paste0(stringr::str_replace_all(sample_metadata$`Factor Value[line]`, " ", "_"), ".fastq")) %>%
  dplyr::select(file_name, new_file_names)

head(simplified_metadata)
```

```{r}
write.table(simplified_metadata, paste0(root, "data/","simplified_metadata.txt"),sep="\t",row.names=FALSE)
```


